---
title: "RestaurantVFC_DataMunging"
author: "Kazım Anıl Eren"
date: "11 Ocak 2018"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### Libraries

```{r Libraries, echo = TRUE}
library(knitr)   # For designing R Markdown document.
library(leaflet) # For geo-graphical analysis.
```

### Data Input 

```{r Data Input, include = FALSE}
rm(list = ls()); gc() # Clearence
air_reserve <- fread(input = "csv_files/air_reserve.csv", encoding = "UTF-8")
air_store   <- fread(input = "csv_files/air_store_info.csv", encoding = "UTF-8") 
air_visit   <- fread(input = "csv_files/air_visit_data.csv", encoding = "UTF-8")
date        <- fread(input = "csv_files/date_info.csv", encoding = "UTF-8")
hpg_reserve <- fread(input = "csv_files/hpg_reserve.csv", encoding = "UTF-8")
hpg_store   <- fread(input = "csv_files/hpg_store_info.csv", encoding = "UTF-8") 
id_relation <- fread(input = "csv_files/store_id_relation.csv", encoding = "UTF-8")
test        <- fread(input = "sample_submission.csv", encoding = "UTF-8")
submission  <- copy(test)
weather     <- fread(input = "japan/weather_data.csv", encoding = "UTF-8")
```

### Data Corrections 

It seems there are some area_names with two distinct co-ordinates.

```{r AIR Coordinates Detection}
kable(air_store[, .N, .(air_area_name, latitude, longitude)][, .N, .(air_area_name)][N > 1])
```

```{r AIR Coordinates Correction}
storelist <- air_store[, .N, .(air_area_name, latitude, longitude)][, .N, .(air_area_name)][N > 1]$air_area_name
storelist <- air_store[air_area_name %in% storelist, .N, .(air_area_name, latitude, longitude)]
storelist[, Real := ifelse(N == max(N), 1,0) , .(air_area_name)]
air_store <- merge(air_store, 
                   storelist[Real == 1, .(air_area_name, coorX = latitude, coorY = longitude)],
                   by = "air_area_name", all.x = T)
air_store[!is.na(coorX) & (latitude != coorX | longitude != coorY),
          ':='(latitude = coorX, longitude = coorY)]
air_store[, ':='(coorX = NULL, coorY = NULL)]
rm(storelist)
```

```{r HPG Coordinates Detection}
kable(hpg_store[, .N, .(hpg_area_name, latitude, longitude)][, .N, .(hpg_area_name)][N > 1])
```

```{r HPG Coordinates Correction}
storelist <- hpg_store[, .N, .(hpg_area_name, latitude, longitude)][, .N, .(hpg_area_name)][N > 1]$hpg_area_name
storelist <- hpg_store[hpg_area_name %in% storelist, .N, .(hpg_area_name, latitude, longitude)]
storelist[, Real := ifelse(N == max(N), 1,0) , .(hpg_area_name)]
storelist <- storelist[hpg_area_name != "None None None"] # One Cannot correct no-location's coordinates.
hpg_store <- merge(hpg_store, 
                   storelist[Real == 1, .(hpg_area_name, coorX = latitude, coorY = longitude)],
                   by = "hpg_area_name", all.x = T)
hpg_store[!is.na(coorX) & (latitude != coorX | longitude != coorY),
          ':='(latitude = coorX, longitude = coorY)]
hpg_store[, ':='(coorX = NULL, coorY = NULL)]
rm(storelist)
```

We can divide area_name column into three and separate date and time info from visit_datetime column.

```{r Separating Area Names & Dates}
air_reserve <- air_reserve[, .(air_store_id,
                               visit_date   = as.Date(str_split_fixed(visit_datetime, pattern = " ", n = 2)[, 1]),
                               visit_hour   = as.numeric(substr(str_split_fixed(visit_datetime, pattern = " ", n = 2)[, 2], 1, 2)),
                               reserve_date = as.Date(str_split_fixed(reserve_datetime, pattern = " ", n = 2)[, 1]),
                               reserve_hour = as.numeric(substr(str_split_fixed(reserve_datetime, pattern = " ", n = 2)[, 2], 1, 2)),
                               visitors     = reserve_visitors
                               )]

air_store[, ':='(air_district1 = str_split_fixed(air_area_name, " ", n = 3)[,1],
                 air_district2 = str_split_fixed(air_area_name, " ", n = 3)[,2],
                 air_district3 = str_split_fixed(air_area_name, " ", n = 3)[,3]
                )]
air_store[, air_area_name := NULL]

air_visit[, visit_date := as.Date(visit_date)]

date        <- date[, .(date = as.Date(calendar_date), 
                        day  = as.factor(day_of_week), 
                        holiday_flg)]

hpg_reserve <- hpg_reserve[, .(hpg_store_id,
                               visit_date   = as.Date(str_split_fixed(visit_datetime, pattern = " ", n = 2)[, 1]),
                               visit_hour   = as.numeric(substr(str_split_fixed(visit_datetime, pattern = " ", n = 2)[, 2], 1, 2)),
                               reserve_date = as.Date(str_split_fixed(reserve_datetime, pattern = " ", n = 2)[, 1]),
                               reserve_hour = as.numeric(substr(str_split_fixed(reserve_datetime, pattern = " ", n = 2)[, 2], 1, 2)),
                               visitors     = reserve_visitors
                              )]

hpg_store   <- hpg_store[, ':='(hpg_district1 = str_split_fixed(hpg_area_name, " ", n = 3)[,1],
                                hpg_district2 = str_split_fixed(hpg_area_name, " ", n = 3)[,2],
                                hpg_district3 = str_split_fixed(hpg_area_name, " ", n = 3)[,3]
                                )]
hpg_store[, hpg_area_name := NULL]

test        <- test[, .(air_store_id = paste0(str_split_fixed(id, "_", n = 3)[,1], "_", str_split_fixed(id, "_", n = 3)[,2]),
                        date   = as.Date(str_split_fixed(id, "_", n = 3)[,3]),
                        visitors)] # visitors column is full of zeros since there has been no forecast all along.
weather     <- weather[, date := as.Date(date, format = "%d.%m.%Y")]
```

There are some stores that match in those datasets. So let's add data from hpg_stores data.table which match with air_stores data.table. Then we'll drop them from hpg_stores data.table to avoid duplication.

```{r AIR & HPG Merge}
air_reserve <- rbind(air_reserve,
                     cbind(merge(hpg_reserve, id_relation, by = "hpg_store_id")[, 7], 
                           merge(hpg_reserve, id_relation, by = "hpg_store_id")[, 2:6]))
hpg_reserve <- hpg_reserve[!hpg_store_id %in% id_relation$hpg_store_id]
```

I will introduce a variable called **PreTime** which will denote that "how many hours before a reservation is made prior to the visit". I also will split visit_hour variable into 4 categorical variables called **night**, **noon**, **afternoon** and **evening** to represent 4 parts of the day.

```{r Introduction of Activity Date}
air_reserve[, PreTime := as.numeric(visit_date - reserve_date) * 24 + visit_hour - reserve_hour]
air_reserve[visit_hour < 8,                      period := "night"]
air_reserve[visit_hour >= 8  & visit_hour <= 13, period := "noon"]
air_reserve[visit_hour >= 14 & visit_hour <= 16, period := "afternoon"]
air_reserve[visit_hour >= 17 & visit_hour <= 21, period := "evening"]
air_reserve[visit_hour > 21,                     period := "night"]

air_daily <- dcast.data.table(air_reserve, visit_date + air_store_id ~ period, fun.aggregate = sum, value.var = "visitors")
air_daily[, reservations := afternoon + evening + night + noon]
air_daily <- merge(air_daily,
                   air_reserve[, .(meanPreTime = mean(PreTime)), .(air_store_id, visit_date)],
                   by = c("air_store_id", "visit_date"), all.x = TRUE)
```

Position or the district of a restaurant is key in determining the number of visitors. Thus, I will create *Amount of Restaurants* and *Amount of Different Genres* variables to calcualte the popularity of a district while *Same Type of Restaurants* variable for determining the competition of that restaurant within its district.

```{r District Analysis}
air_store <- merge(air_store,
                   air_store[, .(AmountofRes = .N, AmountofGenre = length(unique(air_genre_name)))
                             , .(latitude, longitude)],
                   by = c("latitude", "longitude"))
air_store <- merge(air_store,
                   air_store[, .(SameTypeRes = .N), .(latitude, longitude, air_genre_name)],
                   by = c("latitude", "longitude", "air_genre_name"))

leaflet(air_store) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(~longitude, ~latitude, 
             popup = ~air_genre_name,
             clusterOptions = markerClusterOptions())
```

Now let's combine data.sets into one.

```{r Voltran}
air_visit   <- merge(air_visit, 
                     air_daily, 
                     by = c("air_store_id", "visit_date"), all = T)
air_visit   <- merge(air_visit,
                     date[, .(visit_date = date,
                              visit_day  = day, 
                              visit_holiday = holiday_flg)],
                     by = "visit_date", all.x = T)
air_visit[is.na(visitors), visitors := 0]
air_visit[is.na(reservations), ':='(afternoon = 0, evening = 0, night = 0, noon = 0, reservations = 0)]

air_visit   <- merge(air_visit,
                     air_store[, c(2,1,5,6,7)],
                     by = "air_store_id", all.x = T)
air_visit   <- merge(air_visit,
                     weather,
                     by.x = c("air_district1", "visit_date"), by.y = c("air_district1", "date"), all.x = T)
```

```{r Save}
save.image(file = "RV_Forecasting.RData")
```

